---
title: "R Notebook"
output: html_notebook
---

 My code turned into a twisted mess, so I organized the notebook to reproduce the results of my analysis. If you want to dig into why I made the choices I did you'll have to dig into the mess of R files.
 
Read in and clean-up the data:
```{r}
ozone<-read.csv("Ozone.csv")
ozone$Date<-NULL
names(ozone)<-c("Longitude", "Latitude", "Ozone")
cmaq<-read.csv("CMAQ.csv")
cmaq$Date<-NULL
names(cmaq)<-c( "Longitude", "Latitude", "OzonePred")
locs<-read.csv("PredLocs.csv")

```

Match up locations to nearest cmaq predictions:

```{r}
require(rgeos)
ozonePoints <- sp::SpatialPoints(data.frame(ozone$Latitude, ozone$Longitude))
CMAQPoints <- sp::SpatialPoints(data.frame(cmaq$Latitude, cmaq$Longitude))
LOCPoints<-sp::SpatialPoints(data.frame(locs$Latitude, locs$Longitude))

mins<-function(data){
  order(data)[1:100]
}

hundredClosestCmaqRows.ozone <- apply(gDistance(ozonePoints, CMAQPoints, byid=TRUE), 2, mins)
hundredClosestCmaqRows.locs<- apply(gDistance(LOCPoints, CMAQPoints, byid=TRUE), 2, mins)
cmaqScores<-matrix(cmaq$OzonePred[hundredClosestCmaqRows.ozone], byrow = T, nrow=800)
cmaqScoresLoc<-matrix(cmaq$OzonePred[hundredClosestCmaqRows.locs], byrow = T, nrow=nrow(locs))
ozone$cmaqPreds<-cmaqScores
locs$cmaqPreds<-cmaqScoresLoc
ozone$nearestInCmaq<-cmaqScores[,1]

```
Save All data for faster loading next time
```{r}
save(cmaq, CMAQPoints, hundredClosestCmaqRows.ozone, hundredClosestCmaqRows.locs, LOCPoints, locs, ozone, ozonePoints, mins, file="CleanData.RData")
```

Load the data instead of processing (start here):
```{r}
load("CleanData.RData")
```


Center and scale x matrix before performing PCA
```{r}
for (i in 1:100){
  scaling.mean<-mean(ozone$cmaqPreds[,i])
  scaling.sd<-sd(ozone$cmaqPreds[,i])
  ozone$cmaqPreds[,i]<-(ozone$cmaqPreds[,i]-scaling.mean)/scaling.sd
  locs$cmaqPreds[,i]<-(locs$cmaqPreds[,i]-scaling.mean)/scaling.sd
}
```

Compute the principal components of the model
```{r}
psi<- eigen(cov(ozone$cmaqPreds))$vectors[,1:8]
ozone$PCs=ozone$cmaqPreds%*%psi
locs$PCs<-locs$cmaqPreds%*%psi

```

Fit the model using the principal components
```{r}
require(nlme)
model<-gls(Ozone~PCs, correlation = corExp(form=~Latitude+Longitude, nugget = T), data=ozone, method="ML")

```

Extract quantities of interest from model

```{r}
require(LatticeKrig)

sigma2<-model$sigma^2
dist.matrix<-rdist(cbind(ozone$Latitude, ozone$Longitude))
nugget <- coef(model$modelStruct$corStruct,uncons = FALSE, allCoef = TRUE)[2]
range<-coef(model$modelStruct$corStruct,uncons = FALSE, allCoef = TRUE)[1]
R<-sigma2*(((1-nugget)*exp(-dist.matrix/range))+nugget*diag(nrow(ozone)))
```

Compute predictions and residuals from model
```{r}
N=800
K<-800
Xstar<-model.matrix(~PCs, data=ozone)
bhat<-coef(model)
Y<-ozone$Ozone
X<-Xstar
expanded.dist<-rdist(cbind(c(ozone$Latitude, ozone$Latitude), c(ozone$Longitude, ozone$Longitude)))
expanded.R=exp(-expanded.dist/range)
expanded.R<-(1-nugget)*expanded.R+nugget*diag(nrow(expanded.R))
pred.mn<-Xstar%*%bhat + expanded.R[N+(1:K), (1:N)]%*%solve(expanded.R[(1:N),(1:N)])%*%
  (Y-X%*%bhat)

ozone$Prediction<-pred.mn
ozone$Residual<-Y-pred.mn
```

Get Predictions and Prediction variances for unmeasured locations
```{r}
N=800
K<-2685
Xstar<-model.matrix(~PCs, data=locs)
bhat<-coef(model)
Y<-ozone$Ozone
Xmatrix<-model.matrix(~PCs, data=ozone)
expanded.dist<-rdist(cbind(c(ozone$Latitude, locs$Latitude), c(ozone$Longitude, locs$Longitude)))
expanded.R=exp(-expanded.dist/range)
expanded.R<-(1-nugget)*expanded.R+nugget*diag(nrow(expanded.R))
pred.mn<-Xstar%*%bhat + expanded.R[N+(1:K), (1:N)]%*%solve(expanded.R[(1:N),(1:N)])%*%
  (Y-Xmatrix%*%bhat)
pred.var<-diag(sigma2*(expanded.R[N+(1:K),N+(1:K)]-expanded.R[N+(1:K),
                                     (1:N)]%*%solve(expanded.R[(1:N),(1:N)])%*%expanded.R[(1:N),N+(1:K)]))
locs$Prediction<-pred.mn
locs$PredictionVariance<-pred.var
```

Backtransform coefficients to be on original components
```{r}
Z.red<-ozone$PCs
theta.hats<-coef(model)[-1]
beta.hats<-psi%*%theta.hats
var.betas<-diag(psi%*%solve(t(Z.red)%*%solve(R,Z.red), t(psi)))
lowers<-beta.hats-qnorm(.975)*sqrt(var.betas)
uppers<-beta.hats+qnorm(.975)*sqrt(var.betas)
significant<-lowers>0|uppers<0

#require(xtable)
#xtable(data.frame(beta.hats[1:10], CI=ints[1:10]))

```

Uncorrelate model to check assumptions and calculate R^2
```{r}
L<-t(chol(R))
X<-model.matrix(~PCs, data=ozone)
ystar<-solve(L,ozone$Ozone)
xstar<-solve(L, X)
uncor.mod<-lm(ystar~-1+xstar)

```

```{r}
basic.mod<-gls(Ozone~1, correlation = corExp(form=~Latitude+Longitude, nugget = T), data=ozone, method="ML")
anova(model,basic.mod)
summary(basic.mod)
```

